NMRlipids III: Lipid--Cholesterol Interactions in Atomistic Molecular Dynamics Simulations

### 2023-01-03 MANUSCRIPT/manuscript.tex

@@

\begin{abstract}
Cholesterol is a central building block of biomembranes, where it induces orientational order, slows down diffusion, and renders membranes generally stiffer. Molecular dynamics simulations have played a crucial role in resolving these effects and the underlying molecular picture, yet it has recently become evident that different simulation models provide quantitatively different behavior. As the field progresses towards simulations of more complex membranes, the fundamental interactions such as those between cholesterol and phospholipids---the key building blocks of biomembranes---such be modelled accurately.
%
Here, we perform a systematic comparison of the commonly used force fields to describe the structure and dynamics of binary mixtures of phosphatidylcholine and cholesterol. Comparison to deuterium order parameters and lateral diffusion coefficients from NMR as well as electron density profiles from X-ray scattering reveals that none of the existing force field outperforms the others across the physiologically relevant cholesterol concentrations. The Amber-compatible force fields show the smallest deviations from experiment with Slipids and Lipid17 models excelling at low and high cholesterol concentrations, respectively. The development of force fields or the refinement of existing ones can benefit from the included X-ray scattering data set.
\end{abstract}

@@

Cellular membranes contain an incredibly complex mixture of lipid molecules \cite{lorent2020plasma} which are unevenly distributed in the membrane plane and across the leaflets \cite{van2008membrane,wang2020membrane,kinnun2020lateral}. A key player driving the lateral heterogeneity is cholesterol (CHOL), which is present at concentrations between $\sim$10~mol-\% (endoplasmic reticulum) and up to $\sim$50~mol-\% (plasma membrane) \cite{van2008membrane}. CHOL has the unique ability to order neighbouring lipids and thus induce the liquid ordered (L\textsubscript{o}) phase in model membranes \cite{mouritsen2004s,ipsen87,kinnunen91,rog2009ordering}. In the cellular setting, the interaction between phospholipids and CHOL is associated with the formation of lipid rafts and nanodomains \cite{Simons97,cebecauer2018membrane}. This heterogeneity can then further regulate protein distribution \cite{milovanovic2015hydrophobic} or conformation \cite{kelkar2007modulation} \textit{via} the hydrophobic mismatch mechanism \cite{killian1998hydrophobic}, in addition to the direct modulation of protein function by bound CHOL molecules \cite{gimpl2016interaction}.

While detailed experimental information of the interactions among phospholipids and CHOL is relatively sparse, atomistic resolution molecular dynamics (MD) simulations have been widely applied to provide a detailed view of the lateral organization of lipid bilayers as well as lipid--CHOL interactions \cite{rog14,rog2009ordering,berkowitz2009detailed}. However, the trustworthiness of any prediction from an MD simulation relies on the quality of the used model. Fortunately, atomistic \emph{force fields} have seen rapid development recently \cite{leonard2019developing}. The traditional protein force fields CHARMM \cite{brooks1983charmm}, AMBER \cite{cornell1995second}, and OPLS \cite{jorgensen1988opls,harder2016opls3} now have growing libraries of compatible lipid molecules, including CHOL, in the forms of CHARMM36 \cite{Klauda06,lim12}, Lipid17/Slipids \cite{dickson14,madej15,jambeck12,jambeck12b,jambeck13b,grote2020optimization}, and the model by Maciejewski and Rog (here ``MacRog'') \cite{maciejewski14,kulig14,kulig15,Kulig15b}, respectively. Moreover, OPLS3e \cite{harder2016opls3} contains parameters for lipids, but CHOL is currently not included. Thanks to the compatibility with protein force fields, the lipid models can be used in studies of membrane proteins or protein--membrane interactions in addition to sole lipid membranes. Notably, simulations using CHARMM36, Lipid17, and Slipids are now readily set up using CHARMM-GUI for multiple simulation engines \cite{lee16,lee2020charmm}. Despite this ease-of-use, the choice of a force field for an MD simulation can be challenging, as each model and even more so their combinations come with their own sets of pros and cons. Unfortunately, careful comparisons of force field quality, especially for capturing the features of lipid mixtures, are scarce in the literature. This largely results from the lack of systematic, comprehensive, and quantitative datasets from both experiments and simulations.

We have previously performed such comparisons for the structure of phosphatidylcholine (PC) lipids \cite{botan15}, for PC--ion interactions \cite{catte2016molecular}, for phosphatidylserine lipids \cite{antila2019headgroup}, and for phoshatidylethanolamine and phosphatidylglycerol lipids \cite{bacle2021inverse}. Here, we extend this work to mixtures of PC and CHOL. We compare the ability of popular force fields to reproduce effects of CHOL on the structure and dynamics of a PC lipid bilayer, and thereby set guidelines for future efforts to validate intermolecular interactions in binary and more complex systems. Moreover, if some of the studied force fields satisfactorily reproduce experimental behavior at least in some range of CHOL concentrations, the molecular level picture provided by it can be assumed to mimic reality in this concentration regime.

Our earlier work has demonstrated that the structural ensembles of lipids from an MD simulation can be directly compared to those from solid state NMR experiments \cite{botan15,catte2016molecular,antila2019headgroup,bacle2021inverse}. This approach based on order parameters is quantitative, has small error estimates, is highly reproducible, and does not suffer from perturbation of probes used in many other techniques. Thus, we compare our simulation results with different force fields to the systematic NMR data set of PC C--H bond order parameters with varying cholesterol concentrations \cite{ferreira13}. In addition, we extracted electron density profiles across the lipid bilayer from scattering form factor data measured in this work for a range of PC--CHOL mixtures using small angle X-ray scattering. These profiles can be directly compared with the same profiles calculated from the simulation data. Whereas NMR provides data on the conformations of individual lipids, form factor captures the details of the structure of the entire lipid bilayer \cite{ollila16}. One limitation in comparing simulations of lipid mixtures with experimental data is that the various experiments are performed at different concentrations, and thus mimicking all compositions in simulations is tedious. Here, we demonstrate that data extracted from well-sampled simulations can readily be interpolated to any CHOL concentration, allowing the comparison of a single simulation set to experiments performed at different CHOL concentrations. We also evaluated whether the simulation models reproduce the dependence of lateral diffusion coefficients on cholesterol measured using pulsed field gradient NMR \cite{filippov2003effect,filippov2003influence}. This comparison is enabled by the recent theoretical work that allows for the elimination of finite-size effects in MD simulations and thus a quantitative comparison with experiment \cite{vogele2016divergent,vogele2018hydrodynamics}. With the structural and dynamic comparisons established, we estimated the quality of each force field at different CHOL concentrations, identified the best models, and analyzed their molecular organization as we expect them to best reflect reality.

@@

\subsection{X-ray Scattering Experiments}
\todo{Scattering methods incomplete}

Such scattering experiments are challenging to interpret for multi-component lipid bilayers \cite{pan12,Heftberger15,Marquardt15}, yet we still extract 
thus area per molecule values, one of the main quantity used to
compare simulations to experiments, have not been
available for lipid cholesterol mixtures.

SAXS data on POPC multilamellar vesicles (MLVs) at various CHOL concentrations has been measured. Data have been obtained at the EMBL BioSAXS beamline (Hamburg) using 20~keV photons, $T=27^\circ$C. Data were analyzed in terms of the SDP-GAP model described in Refs.~\citenum{Heftberger.2014} and \citenum{Heftberger15}. Data from MLVs are a convolute of structure factor (the crystalline lattice) and form factor. By fitting the scattered intensity data we obtain both contributions. Here we posted only form factors (ASCII format). For information on the quality of the fit we also give plots of the fitted intensity data. The electron density profile has been modelled in terms of the SDP model \cite{heberle12,Kucerka08a,kucerka12}, that is volume distribution functions are modelled by individual Gaussians or error functions. CHOL is challenging to describe within the SDP model, and instead of density profiles, the form factors are often compared between simulation and experiment \cite{Kucerka08b}, or CHOL density is modeled together with that of CH$_2$ groups \cite{pan12}. Here, cholesterol is accounted for by two Gaussians, as proposed by Jianjun Pan (private communication), but is to the best of our knowledge not published.
%(see also PhD Thesis by Peter Heftberger). 
Additional figures show the volume distribution functions and the resulting electron density profiles.

Authors to consult and potentially include in publications using this data: Peter Heftberger (peter.heftberger@gmx.at), Georg Pabst (georg.pabst@uni-graz.at)

@@

We performed MD simulations of a pure POPC membrane as well as five POPC/cholesterol mixtures with the cholesterol content ranging from 11 to 47~mol-\%. In order to eliminate the finite size effects due to periodic boundary conditions from lateral diffusion coefficients of lipids, we performed all simulations in three sizes. The small (64 POPC) and medium (256 POPC) systems were only used in the diffusion analysis, whereas all other analyses were performed on the large membranes (1024 POPC). The number of POPC molecules was kept constant across the different cholesterol concentrations. All membranes were solvated by 50 waters per lipid (POPC or cholesterol). The simulations were performed using four commonly used force fields, namely CHARMM36 \cite{Klauda06,lim12}, Amber-compatible Slipids \cite{jambeck12,jambeck12b,jambeck13b} with its 2020 update \cite{grote2020optimization}, Amber-compatible Lipid17 \cite{dickson14,madej15}, and OPLS-aa-compatible MacRog \cite{kulig14,kulig15,Kulig15b} models. All simulations were 1~\textmu{}s long, totaling 72~\textmu{}s, and performed using GROMACS version 2020 or 2021 \cite{pall2020heterogeneous}. The used simulation parameters are provided in Table~\ref{SItab:params}. The simulation data are available at DOI: 10.5281/zenodo.7035350 (CHARMM36), DOI: 10.5281/zenodo.7022749 (Slipids), DOI: 10.5281/zenodo.6992065 (Lipid17), and DOI: 10.5281/zenodo.7061800 (MacRog).

\begin{table*}[]
\begin{center}
    \caption{\label{tab:simulations}%
    Details of the simulation systems provided for (small/medium/large) system sizes). The box dimensions in the membrane plane ($x/y$) and normal to the membrane ($z$) are provided for the Slipids simulations, and the values vary slightly between the force fields.
    }
    \begin{tabular}{c|ccccc}
    \toprule
    Cholesterol & POPC & Cholesterol & Water & $x/y$ (nm) & $z$ (nm) \\
    \midrule
    0~mol-\%    & 64/256/1024 & 0/0/0       &   3200/12800/51200 & 4.4/9.0/18.1 & 8.9/8.6/8.5    \\
    11~mol-\%   & 64/256/1024 & 8/32/128    &   3600/14400/57600 & 4.5/9.4/18.1 & 9.4/9.1/9.3    \\
    20~mol-\%   & 64/256/1024 & 16/64/256   &   4000/16000/64000 & 4.5/9.2/18.3 & 10.2/9.9/10.0  \\
    29~mol-\%   & 64/256/1024 & 26/104/416  &   4500/18000/72000 & 4.6/9.2/18.5 & 10.8/10.8/10.7 \\
    38~mol-\%   & 64/256/1024 & 40/160/640  &   5200/20800/83200 & 4.8/9.5/19.2 & 11.3/11.4/11.2 \\
    47~mol-\%   & 64/256/1024 & 56/224/896  &   6000/24000/96000 & 5.0/10.1/20.0 & 11.8/11.6/11.7 \\
    \bottomrule
    \end{tabular}
\end{center}
\end{table*}

@@

\paragraph{Electron Density Profiles and Bilayer Dimensions:} The electron density profiles were calculated using \texttt{gmx density} after assigning all atom types with valence electron counts. The thicknesses were extracted by the \texttt{findpeaks} function in Matlab after smoothening the electron density data. Area per phospholipid was obtained by dividing the area of the bilayer by the number of phospholipids in one leaflet.

\paragraph{Lateral diffusion coefficients:} The lateral diffusion coefficients $D_\mathrm{PBC}$ from simulations performed using periodic boundary conditions (PBC) were extracted from mean squared displacement (MSD) data calculated for lipid centers of mass after eliminating the drift of their host leaflet. The MSD data were fit with a straight line in the lag time ($\Delta$) interval between 10 and 100~ns as
%
\begin{align}
	\mathrm{MSD}=4D_\mathrm{PBC}\Delta.
\end{align}

The diffusion coefficients extracted from the three simulation box sizes were fitted with
%
\begin{align}
	D_\mathrm{PBC}\approx D_\infty+\frac{k_\mathrm{B}T}{4\pi\mu_\mathrm{m}h}\frac{\ln\left[L/\left(L_\mathrm{SD}+1.565H\right)\right]-1.713}{1+H/L_\mathrm{SD}},
	\label{eq:sdlatpbc}
\end{align}
%
where $D_\infty$ is the lateral diffusion coefficient in an infinite system, $h$ is the hydrodynamic thickness of the membrane, $k_\mathrm{B}T$ is the Boltzmann constant, $T$ is the temperature, $mu_\mathrm{m}$ and $mu_\mathrm{f}$ are shear viscosities of the membrane and the fluid (water), respectively. $H$ is half the thickness of the water layer and $L_\mathrm{SD}=\frac{h\mu_\mathrm{m}}{2\mu_\mathrm{f}}$ the Saffmann--Delbr\"{u}ck length \cite{vogele2018hydrodynamics}. The inter-leaflet friction coefficient does not appear in Eq.~\eqref{eq:sdlatpbc} as we expect it to be infinite, which was found to be a valid assumption for lipid bilayers \cite{vogele2018hydrodynamics}. The water viscosity value of 0.3228~mPa$\cdot$s was interpolated to 298~K from the values for CHARMM TIP3P in Ref.~\citenum{ong2019temperature} and used for all simulations (CHARMM TIP3P and normal TIP3P differ by $\sim$2--3\%).

@@

\todo{Results: volmaps of chains around cholesterol?}\\
\todo{Results: SI: RDFs of POPC/CHOL around CHOL}\\
\todo{Results: Add missing 30\% data to the scattering plots}\\
\todo{FF minima analysis}

@@

\begin{figure}[htb!]
  \centering
  \includegraphics[width=8.7cm]{../FIGS/dynamics.pdf}
  \caption{\label{fig:dynamics}%
  \textbf{Dynamic properties of the POPC/cholesterol mixtures.}
  Top: Lateral diffusion coefficients corrected for finite-size effects using Eq.~\eqref{eq:dpbc}. Experimental data are taken from NMR measurements on well-hydrated samples \cite{filippov2003influence,filippov2003effect}.
  %
  Bottom: Shear viscosities obtained from the finite-size correction.
  %
  The size-dependence of lateral diffusion and the fits used to obtain the PBC-corrected diffusion coefficients and shear viscosities are shown in Figs.~\ref{SIfig:dvssize} and \ref{SIfig:dvschol}.
  }
\end{figure}

@@

As the results in Fig.~\ref{SIfig:dvschol} demonstrate, the PBC correction can qualitatively change the trends observed for diffusion coefficients. This results from the fact that the size of the PBC correction, Eq.~\eqref{eq:sdlatpbc}, depends on membrane viscosity, which further depends on CHOL concentration. For example, Lipid17 and CHARMM36 underestimate the experimental values in all system sizes used here and the effect of cholesterol seems to be well reproduced. However, the corrected values significantly overshoot the experiment at low CHOL concentration, and CHOL induces a more drastic slowdown in simulations so that the simulations and experiment almost agree at 47~mol-\% CHOL. With MacRog, even the corrected values fall below the experimental ones, yet the slowdown effect of CHOL is again stronger than in experiment. Slipids is a clear outlier, as the CHOL-dependence seems too weak with finite system sizes, yet after accounting for PBC effects, the agreement with experiment is excellent. These results highlight that little can be said about CHOL-dependence of lateral diffusion coefficients without performing multiple simulations required for the PBC correction at multiple CHOL concentrations. When fine-tuning interaction parameters, this can lead to a massive number of simulations.

@@

\begin{figure}[htb!]
  \centering
  \includegraphics[width=\linewidth]{../FIGS/deviation.pdf}
  \caption{\label{fig:deviation}%
  \textbf{Total deviation of the force fields from experimental data.} 
%
  Two leftmost columns:
  The relative deviations from density profiles, palmitate and oleate chain order parameters, and diffusion coefficients are shown in a cumulative manner to highlight the overall deviation of the force fields from experimental data.
  %
  The order parameter deviations are obtained by averaging over the columns in Fig.~\ref{fig:OPmaps} and normalizing against experimental data. The density profile deviations are calculated as the difference between the simulation and experimental maps in Fig.~\ref{fig:densmaps} followed by averaging over the columns and normalization against experiment. The diffusion coefficient deviation is the difference of values from simulation and experiment in Fig.~\ref{fig:dynamics}, taken after interpolation to the same CHOL values as shown in Figs.~\ref{fig:OPmaps} and \ref{fig:densmaps}, and normalized against experimental values.
%
  Top right panel:
  The zoomed-in view to the deviation in density profiles.
%
  Bottom right panel:
  Effect of system size on the deviation of density profiles for the CHARMM36 simulations.
  }
\end{figure}

>>>>

\todo{Update Author list \& Affiliations}

@@

\begin{abstract}
Cholesterol is a central building block in biomembranes, where it induces orientational order, slows down diffusion, and thereby renders the membrane stiffer. Molecular dynamics simulations have played a crucial role in resolving these effects and the underlying molecular picture, yet it has recently become evident that different simulation models predict quantitatively different behavior. Such limitations are easily neglected, as the field rapidly progresses towards simulations of complex membranes mimicking the \textit{in vivo} conditions. Still, for their validity, it is crucial that the interactions between the fundamental building blocks of biomembranes, such as phospholipids and cholesterol, are modelled accurately.
%
Here, we perform a systematic comparison of the ability of commonly used force fields to describe the structure and dynamics of binary mixtures of phosphatidylcholine and cholesterol. Comparison of lipid bilayer simulations with deuterium order parameters and lateral diffusion coefficients from NMR as well as form factors from X-ray scattering reveals that none of the existing force field outperforms the others across the physiologically relevant cholesterol concentrations. The Amber-compatible force fields show the smallest deviations from experiment with Slipids and Lipid17 models excelling at low and high cholesterol concentrations, respectively. 
%The performance of the commonly used CHARMM36 model is subpar, as it overestimates diffusion at low cholesterol concentrations, and gets excessively ordered upon the addition of cholesterol. 
The accompanied X-ray scattering data set measured for this work will foster future force field development and refinement.
\end{abstract}

@@

\todo{Check that the uses of ``model'' and ``force field'' are consistent}

Cellular membranes contain an incredibly complex mixture of lipid molecules \cite{lorent2020plasma} which are unevenly distributed in the membrane plane and across its leaflets \cite{van2008membrane,wang2020membrane,kinnun2020lateral}. A key player driving the lateral heterogeneity is cholesterol (CHOL), which is present at concentrations between $\sim$10~mol-\% (endoplasmic reticulum) and up to $\sim$50~mol-\% (plasma membrane) \cite{van2008membrane}. CHOL has the unique ability to order neighbouring lipids and thus induce the liquid ordered (L\textsubscript{o}) phase in model membranes \cite{mouritsen2004s,ipsen87,kinnunen91,rog2009ordering}. In the cellular setting, the interaction between phospholipids and CHOL is associated with the formation of lipid rafts and nanodomains \cite{Simons97,cebecauer2018membrane}. This heterogeneity can then further regulate protein distribution \cite{milovanovic2015hydrophobic} or conformation \cite{kelkar2007modulation} \textit{via} the hydrophobic mismatch mechanism \cite{killian1998hydrophobic}, in addition to the direct modulation of protein function \cite{gimpl2016interaction,guixa2017membrane}.

While detailed experimental information of the interactions among phospholipids and CHOL is relatively sparse, atomistic resolution molecular dynamics (MD) simulations have been widely applied to provide a detailed view of the lateral organization of lipid bilayers as well as lipid--CHOL interactions \cite{rog14,rog2009ordering,berkowitz2009detailed}. However, the trustworthiness of any prediction from an MD simulation relies on the quality of the used model. Fortunately, in addition to constant refinement of existing parameters, atomistic \emph{force fields} have also seen the expansion to other classes of biomolecules, including lipids, sugars, and nucleic acids \cite{leonard2019developing,marrink2019computational}. The traditional protein force fields CHARMM \cite{brooks1983charmm}, AMBER \cite{cornell1995second}, and OPLS \cite{jorgensen1988opls,harder2016opls3} now have growing libraries of compatible lipid molecules, including CHOL, in the forms of CHARMM36 \cite{Klauda06,lim12}, Lipid17/Slipids \cite{dickson14,madej15,jambeck12,jambeck12b,jambeck13b,grote2020optimization}, and the model by Maciejewski and Rog (here ``MacRog'') \cite{maciejewski14,kulig14,kulig15,Kulig15b}, respectively. Moreover, OPLS3e \cite{harder2016opls3} contains parameters for lipids, but CHOL is currently not included.

Thanks to the compatibility with protein force fields, the lipid models can be used in studies of membrane proteins or protein--membrane interactions in addition to sole lipid membranes. Notably, simulations using CHARMM36, Lipid17, and Slipids are now readily set up using CHARMM-GUI for multiple simulation engines \cite{lee16,lee2020charmm}. Despite this ease-of-use, the choice of a force field for an MD simulation can be challenging, as each model comes with its own set of pros and cons. Unfortunately, careful comparisons of force field quality, especially for capturing the features of lipid mixtures, are scarce in the literature. This largely results from the lack of availability of systematic, comprehensive, and quantitative datasets from both experiments and simulations.

We have previously performed such comparisons for the structure of phosphatidylcholine (PC) lipids \cite{botan15}, for PC--ion interactions \cite{catte2016molecular}, for phosphatidylserine (PS) lipids \cite{antila2019headgroup}, and for phoshatidylethanolamine (PE) and phosphatidylglycerol (PG) lipids \cite{bacle2021inverse}. Here, we extend this work to mixtures of PC and CHOL. We compare the ability of popular force fields to reproduce effects of CHOL on the structure and dynamics of a PC lipid bilayer, and thereby set guidelines for future efforts to validate intermolecular interactions in binary and more complex systems. Moreover, if some of the studied force fields satisfactorily reproduce experimental behavior at least in some range of CHOL concentrations, the molecular level picture provided by it can be assumed to mimic reality in this concentration regime.

Our earlier work has demonstrated that the structural ensembles of lipids from an MD simulation can be directly compared to those from solid state NMR experiments \cite{botan15,catte2016molecular,antila2019headgroup,bacle2021inverse}. This approach based on order parameters is quantitative, has small error estimates, is highly reproducible, and does not suffer from perturbation of probes used in many other techniques. Thus, we compare our simulation results with different force fields to the systematic NMR data set of PC C--H bond order parameters with varying CHOL concentrations \cite{ferreira13}. In addition, we measured form factors for a range of PC--CHOL mixtures using small angle X-ray scattering, and used them to subsequently model the electron density profiles across the lipid bilayer. The form factor and electron density profiles can be compared with the same profiles calculated from the simulation data. Whereas NMR provides data on the conformations of individual lipids, form factors from scattering captures the details of the structure of the entire lipid bilayer \cite{ollila16}.

One limitation in comparing simulations of lipid mixtures with experimental data is that the various experiments are performed at different concentrations, and thus mimicking all compositions in simulations is tedious. Here, we demonstrate that if the simulations and experiments sample multiple CHOL concentrations, the order parameters and electron density profiles can be readily interpolated to intermediate CHOL concentrations. This enables the comparison of data sets that are not extracted at the exact same CHOL concentrations, as is usually the case, especially if multiple techniques are combined in the comparison.

We also evaluated whether the simulation force fields reproduce the dependence of lateral diffusion coefficients on cholesterol measured using pulsed field gradient (PFG) NMR \cite{filippov2003effect,filippov2003influence}. This comparison is enabled by the recent theoretical work that allows for the elimination of finite-size effects in MD simulations and thus a quantitative comparison with experiment \cite{vogele2016divergent,vogele2018hydrodynamics}. With the structural and dynamic comparisons established, we estimated the quality of each force field at different CHOL concentrations. We then analyzed the molecular organization in the best-performing force field at low and high CHOL range, as we expect them to best reflect this organization in reality.

@@

\subsection{X-ray Scattering Experiments}

\todo{Complete scattering methods}

%Such scattering experiments are challenging to interpret for multi-component lipid bilayers \cite{pan12,Heftberger15,Marquardt15}, yet we still extract  thus area per molecule values, one of the main quantity used to compare simulations to experiments, have not been available for lipid cholesterol mixtures.

We measured small angle X-ray scattering (SAXS) data on multilamellar vesicles (MLVs) composed of POPC and CHOL with the latter present at 0--50~mol-\% with 5~mol-\% increments. The data were obtained at the EMBL BioSAXS beamline (Hamburg) using 20~keV photons at $T=300$~K and analyzed in terms of the SDP-GAP model described in Refs.~\citenum{Heftberger.2014} and \citenum{Heftberger15}. The data from MLVs contain the structure factor (the crystalline lattice) and form factor in a convoluted fashion, yet by fitting the scattered intensity data we obtain both contributions. In this work, only the form factors are discussed. The electron density profiles have been modelled in terms of the SDP model \cite{heberle12,Kucerka08a,kucerka12}, \textit{i.e.} volume distribution functions are modelled by individual Gaussians or error functions. CHOL is particularly challenging to describe within the SDP model, and therefore form factors are often compared between simulation and experiment \cite{Kucerka08b} instead of density profiles, or CHOL density is modeled together with that of CH$_2$ groups \cite{pan12}. Here, cholesterol is accounted for by two Gaussians, as proposed by Jianjun Pan (private communication), and used in Refs.~\citenum{belivcka2017high} and \citenum{heftbergerthesis}. 
%Additional figures show the volume distribution functions and the resulting electron density profiles.
%Authors to consult and potentially include in publications using this data: Peter Heftberger (peter.heftberger@gmx.at), Georg Pabst (georg.pabst@uni-graz.at)

@@

We performed MD simulations of a pure POPC membrane as well as five POPC/CHOL mixtures with CHOL content ranging from 11 to 47~mol-\%. In order to eliminate the finite size effects due to periodic boundary conditions from lateral diffusion coefficients of lipids, we performed all simulations in three sizes. The small (64 POPC molecules) and medium (256 POPC molecules) systems were only used in the diffusion analysis, whereas all other analyses were performed on the large membranes (1024 POPC molecules). The number of POPC molecules was kept constant across the different CHOL concentrations. All membranes were solvated by 50 waters per lipid (POPC or CHOL). The simulations were performed using four commonly used force fields, namely CHARMM36 \cite{Klauda06,lim12}, Amber-compatible Slipids \cite{jambeck12,jambeck12b,jambeck13b} with its 2020 update \cite{grote2020optimization}, Amber-compatible Lipid17 \cite{dickson14,madej15}, and OPLS-aa-compatible MacRog \cite{kulig14,kulig15,Kulig15b} models. All simulations were 1~\textmu{}s long, totaling 72~\textmu{}s, and performed using GROMACS version 2020 or 2021 \cite{pall2020heterogeneous}. The used simulation parameters are provided in Table~\ref{SItab:params}, and the simulation data are available at DOI: 10.5281/zenodo.7035350 (CHARMM36), DOI: 10.5281/zenodo.7022749 (Slipids), DOI: 10.5281/zenodo.6992065 (Lipid17), and DOI: 10.5281/zenodo.7061800 (MacRog).

\todo{Run Lipid21 for revision?}

\begin{table*}[]
\begin{center}
    \caption{\label{tab:simulations}%
    Details of the simulation systems provided for (small/medium/large) system sizes). The box dimensions in the membrane plane ($x/y$) and normal to the membrane ($z$) are provided for the Slipids simulations, and the values vary slightly between the force fields.
    }
    \begin{tabular}{c|ccccc}
    \toprule
    $[$CHOL$]$ & POPC & CHOL & Water & $x/y$ (nm) & $z$ (nm) \\
    \midrule
    0~mol-\%    & 64/256/1024 & 0/0/0       &   3200/12800/51200 & 4.4/9.0/18.1 & 8.9/8.6/8.5    \\
    11~mol-\%   & 64/256/1024 & 8/32/128    &   3600/14400/57600 & 4.5/9.4/18.1 & 9.4/9.1/9.3    \\
    20~mol-\%   & 64/256/1024 & 16/64/256   &   4000/16000/64000 & 4.5/9.2/18.3 & 10.2/9.9/10.0  \\
    29~mol-\%   & 64/256/1024 & 26/104/416  &   4500/18000/72000 & 4.6/9.2/18.5 & 10.8/10.8/10.7 \\
    38~mol-\%   & 64/256/1024 & 40/160/640  &   5200/20800/83200 & 4.8/9.5/19.2 & 11.3/11.4/11.2 \\
    47~mol-\%   & 64/256/1024 & 56/224/896  &   6000/24000/96000 & 5.0/10.1/20.0 & 11.8/11.6/11.7 \\
    \bottomrule
    \end{tabular}
\end{center}
\end{table*}

@@

\paragraph{Electron Density Profiles and Bilayer Dimensions:} The form factors and electron density profiles were calculated using the \texttt{form\_factor.py} script in the NMRlipids Databank (\url{https://github.com/NMRLipids/Databank/}) after assigning all atom types with valence electron counts. The thicknesses were extracted by the \texttt{findpeaks} function in Matlab after smoothening the electron density data. Area per phospholipid was obtained by dividing the area of the bilayer by the number of phospholipids in one leaflet.

\paragraph{Order Parameters:} The C--H order parameters were calculated as
%
\begin{align}
    S_\mathrm{CD}=\left\langle\frac{3\cos^2\theta-1}{2}\right\rangle,
\end{align}
%
where $\theta$ is the angle between the C--H vector and the $z$ axis normal to the bilayer plane. In the POPC acyl chains, the values for the 2 (3) hydrogens of CH$_2$ (CH$_3$) groups were averaged, since these groups rotate freely and thus the order parameters are identical for both (all) hydrogens in experiments. In simulations, the values are also very similar, justifying this approach. An exception to this are the hydrogens bound to the 2\textsuperscript{nd} carbon in the oleate chain; they lack rotational averaging in both both simulations and experiments, and were thus treated separately in our analyses. The C--H order parameters for the POPC head group were calculated similarly, but values for both hydrogens are shown separately for all CH$_2$ groups, \textit{i.e.} no averaging was performed.

\paragraph{Lateral Diffusion Coefficients:} The lateral diffusion coefficients $D_\mathrm{PBC}$ from simulations performed using periodic boundary conditions (PBC) were extracted from mean squared displacement (MSD) data calculated for lipid centers of mass after eliminating the drift of their host leaflet with the \texttt{gmx msd} tool. The MSD data were fit with a straight line in the lag time ($\Delta$) interval between 10 and 100~ns as
%
\begin{align}
	\mathrm{MSD}=4D_\mathrm{PBC}\Delta.
\end{align}

The diffusion coefficients extracted from the three simulation box sizes were fitted with
%
\begin{align}
	D_\mathrm{PBC}\approx D_\infty+\frac{k_\mathrm{B}T}{4\pi\mu_\mathrm{m}h}\frac{\ln\left[L/\left(L_\mathrm{SD}+1.565H\right)\right]-1.713}{1+H/L_\mathrm{SD}},
	\label{eq:sdlatpbc}
\end{align}
%
where $D_\infty$ is the lateral diffusion coefficient in an infinite system, $h$ is the hydrodynamic thickness of the membrane, $k_\mathrm{B}T$ is the Boltzmann constant, $T$ is the temperature, $mu_\mathrm{m}$ and $mu_\mathrm{f}$ are shear viscosities of the membrane and the fluid (water), respectively. $H$ is half the thickness of the water layer and $L_\mathrm{SD}=\frac{h\mu_\mathrm{m}}{2\mu_\mathrm{f}}$ the Saffmann--Delbr\"{u}ck length \cite{vogele2018hydrodynamics}. The inter-leaflet friction coefficient does not appear in Eq.~\eqref{eq:sdlatpbc} as we expect it to be infinite, which was found to be a valid assumption for lipid bilayers \cite{vogele2018hydrodynamics}. The water viscosity value of 0.3228~mPa$\cdot$s was interpolated to 298~K from the values for CHARMM TIP3P in Ref.~\citenum{ong2019temperature} and used for all simulations (CHARMM TIP3P and normal TIP3P differ by $\sim$2--3\%). The final simulation box dimensions were used for $L$ and $H$. For $h$ and $H$ ($=(L_z-h)/2$, the average values from the three systems sizes were used.

@@

\todo{Results: volmaps of chains around cholesterol?}\\
\todo{Results: SI: RDFs of POPC/CHOL around CHOL}\\
\todo{FF minima analysis}

@@

\begin{figure}[htb!]
  \centering
  \includegraphics[width=8.7cm]{../FIGS/dynamics.pdf}
  \caption{\label{fig:dynamics}%
  \textbf{Dynamic properties of the POPC/cholesterol mixtures.}
  Top: Lateral diffusion coefficients corrected for finite-size effects using Eq.~\eqref{eq:dpbc}. Experimental data are taken from NMR measurements on well-hydrated samples \cite{filippov2003influence,filippov2003effect}. The hollow circles show data extracted for Slipids using a shorter Lennard-Jones cutoff (see text).
  %
  Bottom: Shear viscosities obtained from the finite-size correction.
  %
  The size-dependence of lateral diffusion and the fits used to obtain the PBC-corrected diffusion coefficients and shear viscosities are shown in Figs.~\ref{SIfig:dvssize} and \ref{SIfig:dvschol}.
  }
\end{figure}

@@

Slipids has the largest Lennard-Jones (LJ) cutoff of the used force fields of 1.4~nm (see Table~\ref{SItab:params}), and this cutoff is known to affect dynamic properties \cite{leonard2018comparison}. Therefore, we repeated simulations at 0~mol-\% and 47~mol-\% CHOL for Slipids but with a cutoff of 0.9~nm corresponding to that of Lipid17, and the PBC-corrected diffusion coefficient values are shown in the top panel of Fig.~\ref{fig:dynamics} as hollow green circles. The decrease in cutoff indeed leads to a minor increase in the lateral diffusion coefficient. However, the LJ cutoff alone does not explain the differences between force fields, indicating that they truly spark from differences in the force field interaction parameters.

As the results in Fig.~\ref{SIfig:dvschol} demonstrate, the PBC correction can qualitatively change the trends observed for diffusion coefficients. This results from the fact that the size of the PBC correction, Eq.~\eqref{eq:sdlatpbc}, depends on membrane viscosity, which further depends on CHOL concentration. For example, Lipid17 and CHARMM36 underestimate the experimental values in all system sizes used here and the effect of CHOL seems to be well reproduced. However, the corrected values significantly overshoot the experiment at low CHOL concentration, and CHOL induces a more drastic slowdown in simulations so that the simulations and experiment almost agree at 47~mol-\% CHOL. With MacRog, even the corrected values fall below the experimental ones, yet the slowdown effect of CHOL is again stronger than in experiment. Slipids is a clear outlier, as the CHOL-dependence seems too weak with finite system sizes, yet after accounting for PBC effects, the agreement with experiment is excellent. These results highlight that little can be said about CHOL-dependence of lateral diffusion coefficients without performing multiple simulations required for the PBC correction at multiple CHOL concentrations. When fine-tuning interaction parameters, this can lead to a massive number of simulations.

@@

\begin{figure}[htb!]
  \centering
  \includegraphics[width=\linewidth]{../FIGS/deviation.pdf}
  \caption{\label{fig:deviation}%
  \textbf{Total deviation of the force fields from experimental data.} 
%
  Two leftmost columns:
  The relative deviations from the first two minima in the form factor), palmitate and oleate chain order parameters, and diffusion coefficients are shown in a cumulative manner to highlight the overall deviation of the force fields from experimental data.
  %
  The order parameter deviations are obtained by averaging over the columns in Fig.~\ref{fig:OPmaps} and normalizing against experimental data. The differences in the form factor minima in Fig.~\ref{fig:SIfig:ffminima} between experiment and simulation were calculated, normalized against experimental data, and summed together. The diffusion coefficient deviation is the difference of values from simulation and experiment in Fig.~\ref{fig:dynamics}, taken after interpolation to the same CHOL values as shown in Figs.~\ref{fig:OPmaps} and \ref{fig:densmaps}, and normalized against experimental values.
%
  Top right panel: 
  The zoomed-in view to the deviation in the location of the first minimum in the form factor.
%
  Bottom right panel:
  Deviation of the density profiles calculated from the form factor using the SDS model. The density profile deviations are calculated as the difference between the simulation and experimental maps in Fig.~\ref{fig:densmaps} followed by averaging over the columns and normalization against experiment.
  }
\end{figure}
